---
title: "Monte Carlo Simulation of Portfolio Returns for Different Optimization Strategies"
authors: "Madi Kassymbekov; Abderezzak Amimer; Stephanie Vu"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("Simulation.R")
simulations = 1000
tradingdays = 253
```

## Context

The main goal of investors when creating an asset portfolio is to obtain the most value from their constructed portfolio. A major decision in portfolio management lies in defining how to allocate funds when constructing a portfolio. Portfolio optimization is a phenomenon widely studied in finance. It consists of determining the optimal proportions of total capital invested to assign to each particular asset in the portfolio. Such a problem poses a challenge to financial investors as portfolio managers seek to define the best way to distribute capital in order to yield the most favorable portfolio returns depending on the level of risk the investor is willing to take. Many different strategies exist to define optimal weights in a portfolio however, it remains hard for portfolio managers to decide which optimization strategy is best suited for a given set of risky assets.


## Scenario 1: All Normally Distributed Stock Returns

```{r, echo=FALSE, warning=FALSE}
scenario = "allnormal"
set.seed(123)
portfolio_returns = portfolio_simulation(simulations, tradingdays, scenario, "1")
portfolio_returns[[3]]
```

```{r, echo=FALSE, warning=FALSE}
allnorm_ret=na.omit(portfolio_returns[[1]])
allnorm_mean=round(colMeans(na.omit(portfolio_returns[[1]])),6)
  a1<- ggplot(allnorm_ret, aes(x=meanvarReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(meanvarReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Mean Variance",x="Return", y = "Frequency")+
    theme_classic()
  
  a2<- ggplot(allnorm_ret, aes(x=minvarReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(minvarReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Min Variance",x="Return", y = "Frequency")+
    theme_classic()
  
  a3<- ggplot(allnorm_ret, aes(x=maxdivReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(maxdivReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Max Diversification",x="Return", y = "Frequency")+
    theme_classic()
  
  a4<-ggplot(allnorm_ret, aes(x=maxdecReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(maxdecReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Max Decorrelation",x="Return", y = "Frequency")+
    theme_classic()
  
  a5<-ggplot(allnorm_ret, aes(x=equalweightsReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(equalweightsReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Equal Weights",x="Return", y = "Frequency")+
    theme_classic()
  
  
  figure1<-ggarrange(a1,a2,a3,a4,a5,
                     ncol=3,nrow=2)
  
  annotate_figure(figure1,
                  top = text_grob("Portfolio of All Normal Stocks", color = "black", face = "bold", size = 14),
                  fig.lab = "Figure 1", fig.lab.face = "bold")
```

## Scenario 2: All Normally Distributed Stock Returns with low volatility

```{r, echo=FALSE, warning=FALSE}
scenario = "low volatility"
portfolio_returns = portfolio_simulation(simulations, tradingdays, scenario, "3")
portfolio_returns[[3]]
```

```{r, echo=FALSE, warning=FALSE}
lowvol_ret=na.omit(portfolio_returns[[1]])
lowvol_mean=round(colMeans(na.omit(portfolio_returns[[1]])),6)
 c1<- ggplot(lowvol_ret, aes(x=meanvarReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(meanvarReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Mean Variance",x="Return", y = "Frequency")+
    theme_classic()
  
  c2<- ggplot(lowvol_ret, aes(x=minvarReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(minvarReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Min Variance",x="Return", y = "Frequency")+
    theme_classic()
  
  c3<- ggplot(lowvol_ret, aes(x=maxdivReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(maxdivReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Max Diversification",x="Return", y = "Frequency")+
    theme_classic()
  
  c4<-ggplot(lowvol_ret, aes(x=maxdecReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(maxdecReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Max Decorrelation",x="Return", y = "Frequency")+
    theme_classic()
  
  c5<-ggplot(lowvol_ret, aes(x=equalweightsReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(equalweightsReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Equal Weights",x="Return", y = "Frequency")+
    theme_classic()
  
  
  figure3<-ggarrange(c1,c2,c3,c4,c5,
                     ncol=3,nrow=2)
  
  annotate_figure(figure3,
                  top = text_grob("Portfolio of Low Volatility Stocks", color = "black", face = "bold", size = 14),
                  fig.lab = "Figure 3", fig.lab.face = "bold")
```

## Scenario 3: All Normally Distributed Stock Returns with high volatility

```{r, echo=FALSE, warning=FALSE}
scenario = "high volatility"
portfolio_returns = portfolio_simulation(simulations, tradingdays, scenario, "4")
portfolio_returns[[3]]
```

```{r, echo=FALSE, warning=FALSE}
highvol_ret=na.omit(portfolio_returns[[1]])
highvol_mean=round(colMeans(na.omit(portfolio_returns[[1]])),6)
  d1<- ggplot(highvol_ret, aes(x=meanvarReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(meanvarReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Mean Variance",x="Return", y = "Frequency")+
    theme_classic()
  
  d2<- ggplot(highvol_ret, aes(x=minvarReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(minvarReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Min Variance",x="Return", y = "Frequency")+
    theme_classic()
  
  d3<- ggplot(highvol_ret, aes(x=maxdivReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(maxdivReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Max Diversification",x="Return", y = "Frequency")+
    theme_classic()
  
  d4<-ggplot(highvol_ret, aes(x=maxdecReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(maxdecReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Max Decorrelation",x="Return", y = "Frequency")+
    theme_classic()
  
  d5<-ggplot(highvol_ret, aes(x=equalweightsReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(equalweightsReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Equal Weights",x="Return", y = "Frequency")+
    theme_classic()


figure4<-ggarrange(d1,d2,d3,d4,d5,
                   ncol=3,nrow=2)

annotate_figure(figure4,
                top = text_grob("Portfolio of High Volatility Stocks", color = "black", face = "bold", size = 14),
                fig.lab = "Figure 4 ", fig.lab.face = "bold")
```

## Scenario 4: Mixed distributed Stock Returns

```{r, echo=FALSE, warning=FALSE}
scenario = "mix"
portfolio_returns = portfolio_simulation(simulations, tradingdays, scenario, "2")
portfolio_returns[[3]]
```

```{r, echo=FALSE, warning=FALSE}
mix_ret=na.omit(portfolio_returns[[1]])
mix_ret_mean=round(colMeans(na.omit(portfolio_returns[[1]])),6)
   b1<- ggplot(mix_ret, aes(x=meanvarReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(meanvarReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Mean Variance",x="Return", y = "Frequency")+
    theme_classic()
  
  b2<- ggplot(mix_ret, aes(x=minvarReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(minvarReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Min Variance",x="Return", y = "Frequency")+
    theme_classic()
  
  b3<- ggplot(mix_ret, aes(x=maxdivReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(maxdivReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Max Diversification",x="Return", y = "Frequency")+
    theme_classic()
  
  b4<-ggplot(mix_ret, aes(x=maxdecReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(maxdecReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Max Decorrelation",x="Return", y = "Frequency")+
    theme_classic()
  
  b5<-ggplot(mix_ret, aes(x=equalweightsReturn)) +
    geom_histogram(fill="white", color="black", bins=20)+
    geom_vline(aes(xintercept=mean(equalweightsReturn,na.rm=TRUE)), color="blue",
               linetype="dashed")+
    labs(title="Equal Weights",x="Return", y = "Frequency")+
    theme_classic()
  
  
  figure2<-ggarrange(b1,b2,b3,b4,b5,
                     ncol=3,nrow=2)
  
  annotate_figure(figure2,
                  top = text_grob("Portfolio of Mixed Distribution Stocks", color = "black", face = "bold", size = 14),
                  fig.lab = "Figure 2", fig.lab.face = "bold")
```

## Comparison of Portfolio Strategies 

```{r, echo=FALSE, warning=FALSE}
#Confidence Intervals
  #All Normal
  allnorm_CI=MeanCI(allnorm_ret$meanvarReturn, conf.level = 0.95,na.rm=TRUE)
  allnorm_CI=rbind(allnorm_CI,MeanCI(allnorm_ret$minvarReturn, conf.level = 0.95,na.rm=TRUE),
 MeanCI(allnorm_ret$maxdivReturn, conf.level = 0.95,na.rm=TRUE),
 MeanCI(allnorm_ret$maxdecReturn, conf.level = 0.95,na.rm=TRUE),
 MeanCI(allnorm_ret$equalweightsReturn, conf.level = 0.95,na.rm=TRUE))
  allnorm_CI=round(allnorm_CI,6)

  #Mixed
  mix_CI=MeanCI(mix_ret$meanvarReturn, conf.level = 0.95,na.rm=TRUE)
  mix_CI=rbind(mix_CI,MeanCI(mix_ret$minvarReturn, conf.level = 0.95,na.rm=TRUE),
  MeanCI(mix_ret$maxdivReturn, conf.level = 0.95,na.rm=TRUE),
  MeanCI(mix_ret$maxdecReturn, conf.level = 0.95,na.rm=TRUE),
  MeanCI(mix_ret$equalweightsReturn, conf.level = 0.95,na.rm=TRUE))
  mix_CI=round(mix_CI,6)
  
  #Low Volatility
  low_CI=MeanCI(lowvol_ret$meanvarReturn, conf.level = 0.95,na.rm=TRUE)
  low_CI=rbind(low_CI,MeanCI(lowvol_ret$minvarReturn, conf.level = 0.95,na.rm=TRUE),
  MeanCI(lowvol_ret$maxdivReturn, conf.level = 0.95,na.rm=TRUE),
  MeanCI(lowvol_ret$maxdecReturn, conf.level = 0.95,na.rm=TRUE),
  MeanCI(lowvol_ret$equalweightsReturn, conf.level = 0.95,na.rm=TRUE))
  low_CI=round(low_CI,6)
  
  #High Volatility
  high_CI=MeanCI(highvol_ret$meanvarReturn, conf.level = 0.95,na.rm=TRUE)
  high_CI=rbind(high_CI,MeanCI(highvol_ret$minvarReturn, conf.level = 0.95,na.rm=TRUE),
  MeanCI(highvol_ret$maxdivReturn, conf.level = 0.95,na.rm=TRUE),
  MeanCI(highvol_ret$maxdecReturn, conf.level = 0.95,na.rm=TRUE),
  MeanCI(highvol_ret$equalweightsReturn, conf.level = 0.95,na.rm=TRUE))
  high_CI=round(high_CI,6)
  
#Tables 
  #All norm
  t1=data.frame(rbind(allnorm_mean[1:3],allnorm_mean[4:6],allnorm_mean[7:9],allnorm_mean[10:12],allnorm_mean[13:15]))
  t1=cbind(t1,allnorm_CI[,2:3])
  rownames(t1)<-c("Mean Variance","Min Variance","Max Diversification","Max Decorrelation","Equal Weights")
  colnames(t1)<- c("Return (%)","Variance (%)","Sharpe", "Lower CI","Upper CI")
  t1[1:2]<-t1[1:2]*100
  t1[4:5]<-t1[4:5]*100
  t1 %>%
    kbl(caption = "Table 1: Portfolio of All Normal Stocks") %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  #Mixed
  t2=data.frame(rbind(mix_ret_mean[1:3],mix_ret_mean[4:6],mix_ret_mean[7:9],mix_ret_mean[10:12],mix_ret_mean[13:15]))
  t2=cbind(t2,mix_CI[,2:3])
  rownames(t2)<-c("Mean Variance","Min Variance","Max Diversification","Max Decorrelation","Equal Weights")
  colnames(t2)<- c("Return (%)","Variance (%)","Sharpe", "Lower CI","Upper CI")
  t2[1:2]<-t2[1:2]*100
  t2[4:5]<-t2[4:5]*100
  t2 %>%
    kbl(caption = "Table 2: Portfolio of Mixed Distribution Stocks") %>%
    kable_classic(full_width = F, html_font = "Cambria")
  
  #Low Volatility
 t3=data.frame(rbind(lowvol_mean[1:3],lowvol_mean[4:6],lowvol_mean[7:9],lowvol_mean[10:12],lowvol_mean[13:15]))
 t3=cbind(t3,low_CI[,2:3])
  rownames(t3)<-c("Mean Variance","Min Variance","Max Diversification","Max Decorrelation","Equal Weights")
  colnames(t3)<- c("Return (%)","Variance (%)","Sharpe", "Lower CI","Upper CI")
  t3[1:2]<-t3[1:2]*100
  t3[4:5]<-t3[4:5]*100
 t3 %>%
   kbl(caption = "Table 3: Portfolio of Low Volatility Stocks") %>%
   kable_classic(full_width = F, html_font = "Cambria")
 
 #High Volatility
 t4=data.frame(rbind(highvol_mean[1:3],highvol_mean[4:6],highvol_mean[7:9],highvol_mean[10:12],highvol_mean[13:15]))
 t4=cbind(t4,high_CI[,2:3])
 rownames(t4)<-c("Mean Variance","Min Variance","Max Diversification","Max Decorrelation","Equal Weights")
 colnames(t4)<- c("Return (%)","Variance (%)","Sharpe", "Lower CI","Upper CI")
 t4[1:2]<-t4[1:2]*100
 t4[4:5]<-t4[4:5]*100
 t4 %>%
   kbl(caption = "Table 4: Portfolio of High Volatility Stocks") %>%
   kable_classic(full_width = F, html_font = "Cambria")
```